### 2021-07-13 untitled-3.tex
work

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

\usepackage{amsmath,amsthm,amssymb}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{enumerate}
%\usepackage{caption,subcaption}
\usepackage{caption}
%\usepackage{subcaption}

%\usepackage{newpxtext,newpxmath}
\usepackage{multicol}

%\theoremstyle{definition}
\newtheorem{thm}{Theorem}
\newtheorem{cor}{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\theoremstyle{definition}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ex}[thm]{Example}

\numberwithin{equation}{section}
\numberwithin{thm}{section}

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

\newcommand{\winva}{{w_1^{-1}(a)}}
\newcommand{\winvb}{{w_1^{-1}(b)}}
\renewcommand{\a}{a}
\renewcommand{\b}{b}
\newcommand{\m}{n_1}
\newcommand{\mtwo}{n_2}

%%% END Article customizations

%%% The "real" document content comes below...

\title{Brief Article}
\author{The Author}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

Let $\chi_B$ be the characteristic function on $\mathbb{R}$ given by $\chi_B(u)=1$ if $u\in B$ and $\chi_B(u)=0$ otherwise. Assume that $k$ is a bounded and periodic function, and 
$g$ is integrable with period $\phi(>0)$ and is differentiable at least $q-1$ times. In this paper we consider the following assumptions:

\begin{enumerate}
\item $k$ has sufficient decay;
\item $\b-\a>\phi(g)$ and $\b-\a<\frac{\phi(g)\phi(\chi_{[\a,\b]})}{\phi(g)-\phi(\chi_{[\a,\b]})};$
\item $\int^{\infty}_{-\infty}k(x-y)\chi_{[\a,\b]}(y)\,\mathrm{d}y<\int^{\infty}_{-\infty}k(x-y)\chi_{[\b,\infty)}(y)\,\mathrm{d}y$.
\end{enumerate}

The main results of this paper are the following theorems:

\begin{enumerate}
\item The condition $(i)$ and $(iii)$ imply that $k$ is symmetric with respect to $\a$ and $\b$.
\item If $\a-\b>0$, then $g$ is nondecreasing.
\item $u_c=\frac{2\a\b}{\a+\b}$ is the spreading speed for  $g=k+\a-\b$ and $\a+\b>0$. 
$u_c$ is the asymptotic velocity of a spreading mode, and is well defined if the integral of $g$ is finite and positive.
\item There is a unique spreading mode if $\a-\b>0$ and $\int^{\infty}_{-\infty}g(x)\,\mathrm{d}x$ is finite.
\item If $\a-\b<0$, then $g$ is nonmonotone and $g$ has at most one minimum
\end{enumerate}
\end{document}